# Connections

```{r knit-setting-for-con, eval = TRUE, include=FALSE}

knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = TRUE, warning = FALSE, collapse = TRUE)
```

## Issue a connection variable {#issue-con}

- Retruns a new `DBIConnection` object used to communicate with a database.
- It is recommended that you make sure you have disconnected old connections before using this.

### Definition


```{r issue-con-show, ref.label='def-issue-con', eval=FALSE}
```

### Example

Before setting a new connection variable, make sure that you are not trying to issue more connection variables than you need or overwrite exsisting one without disconnect it. So it is a good practice to disconnect before connecting, for instance, using `release_cons` defined below.

```{r issue-con-example}
release_cons() # Don't forget!
con <- issue_con("demo")
con %>% print()
```

## Disconnect connections {#release-cons}

### Definition

- Disconnects all connection variables whose names are included in the passed vector of strings.
- Retruns a vector of characters listing connection variables that are disconnected by the function call.
- The arguments `search_from = ` specifies a vector of strings that might contain connection variables to be disconnected. The default is set to a vector of names of all objects in the global namespace.
- The arguments `except = NULL` specifies a vector of strings that should be excluded from `search_from` argument.  

```{r ref.label='def-release-con', eval=FALSE}
```

### Example

```{r release-cons-example, warning=TRUE}
release_cons() %>% invisible()

con1 <- issue_con("demo")
con2 <- issue_con("demo")
cons <- release_cons()

cons %>% print()

# check that they are actually disconnected
purrr::map_lgl(cons, ~ dbDisconnect(.x %>% get()))

# disconnect all connetions but 'con'
con <- issue_con("demo")
con1 <- issue_con("demo")
con2 <- issue_con("demo")
release_cons(except = "con")
con %>% print() # con is still valid
```

