## Connecting and Disconnecting

### `dbCanConnect()`

- Checks if a connection to a DBMS can be established
- Retruns logical.
- If `FALSE`, it comes with a reason for failure.

```{r can-connect, collapse=TRUE}
DBI::dbCanConnect(drv = RPostgres::Postgres())
RPostgres::dbCanConnect(drv = RPostgres::Postgres())
```

### `dbConnect()` {#dbConnect}

- Creates a connection to a DBMS.
- Retruns S4 object (that inherits from DBIConnection virtual class), which holds the established connection information used to communicate with database.
- Many connection variable can exist at the same time without affecting each other.

The following example shows a typical argument list for connecting to Postgres.

```{r, eval=FALSE}
con <- DBI::dbConnect(
    drv = RPostgres::Postgres(),
    host = "localhost",
    port = 5432,
    dbname = "your-database-name",
    user = "your-user",
    password = "your-password"
)
```

Be aware that this is a terrible way of connecting to a non-demo database because your credentials are completely revealed in a plain text. See [Securing Credentials](https://db.rstudio.com/best-practices/managing-credentials) for some better ways. In this document, we issue connection variables by calling `issue_con` defined in \@ref(Functions-dbi).

```{r, collapse=TRUE}
con <- issue_con("demo")
# connection variable is printable
con %>% print()
```

### `dbDisconnect()`

- Closes the associated connection, discards all associated pending work, and frees resources.
- Retruns `TRUE` invisibly.
- Issues warning on an already disconnected connnectinon variable or invalid one.
- NOT clears a supplied connection object.

```{r disconnect-example, collapse=TRUE, warning=TRUE}
# disconnect con varibale. returns nothing.
con %>% dbDisconnect()
# get warned since con is already disconnected
con %>% dbDisconnect()
```

Disconnected object continues to exist as a non-null object.

```{r disconnected-alive, collapse=TRUE}
con %>% is.null()
```

But it is no longer printable.

```{r disconnected-con-unpintable, collapse=TRUE, error=TRUE}
con %>% print()
```

Disconnection has a local effect. That is, disconnecting one connection has no impact on other connections.

```{r ,collapse=TRUE}
con <- issue_con("demo")
con1 <- issue_con("demo")
con1 %>% dbDisconnect()
# con is still valid
con %>% print()
```

### `dbGetInfo()`

- Retrieves information on database objects, i.e., DBIDriver (e.g., `RPostgres::Postgres`), DBIConnection (e.g., `con`) or DBIResult (e.g., `dbSendQuery()`).
- Returns a named list whose components depend on the input class.

```{r dbGetInfo, collapse=TRUE}
dbGetInfo(RPostgres::Postgres())
con %>% dbGetInfo()
```

### `dbIsReadOnly()`

- Tests whether a database object is read only.
- Returns logical.

```{r dbIsReadOnly, collapse=TRUE}
con %>% dbIsReadOnly()
dbIsReadOnly(RPostgres::Postgres())
```

```{r, include=FALSE}
con %>% dbDisconnect()
```