## Manipulating Tables

```{r ,include=FALSE}
con <- refresh_con(con)
remove_all_table(con)
```

### `dbWriteTable()`

- Writes, overwrites or appends a data frame to a database table
- Various optional argumets (must be provided as named arguments) available such as  
	- `temoprary = FALSE`: a logical specifying whether a created table should be temoprary in the sense that it vanishes after disconection. **NOT available with postgres**.
	- `overwrite = FALSE`: a logical specifying whether to overwrite an existing table.
	- `append = FALSE`: a logical specifying whether to append to an existing table.

The following example demonstrates that `DBI::dbWriteTable()` does not accept `temoprary = ` option. Note that no error or warnig is raised. The same is true for `Postgres::dbWriteTable()`.

```{r write-mtcars-postgres, collapse=TRUE}
# temprary = TRUE is available?
con %>% RPostgres::dbWriteTable("mtcars", mtcars, overwrite = TRUE, temoprary = TRUE)
# show list of tables currently available
con %>% dbListTables()
# refresh connnection
con %>% dbDisconnect()
con <- issue_con("demo")
# is the table still alive? -> yes
con %>% dbListTables()
```

### `dbListTables()`

- Returns a character vector that enumerates all tables and views in the database.
- In Postgres, this includes views and temprary objects created by `dplyr::copy_to()`, but not those by `dbplyr::copy_inline()`.

```{r ListTable, collapse=TRUE}
con %>% dbListTables()
```

### `dbListObjects()`

- Returns a data frame with columns table and is_prefix in that order.
- Works also for temprary tables created by `dplyr::copy_to()` in Postgres.

```{r ListObj, collapse=TRUE}
con %>% dbListObjects()
```

### `dbListFields()`

- Returns a character vector that enumerates all fields in the table in the correct order.
- Works also for temprary tables created by `dplyr::copy_to()` in Postgres.

```{r Listfields, collapse=TRUE}
con %>% dbListFields("mtcars")
```

### `dbExistsTable()`

- Returns a lo'gical scalar, `TRUE` if the table or view specified by the name argument exists, `FALSE` otherwise.
- Works also for temprary tables created by `dplyr::copy_to()` in Postgres.

```{r exsits-table, collapse=TRUE}
con %>% dbExistsTable("mtcars")
con %>% dbExistsTable("hoge")
```

### `dbRemoveTable()`

- Removes a remote table from the database.
- Returns `TRUE` invisibly.
- Raises an error if the intended table does not exist (see `fail_if_missing` option below).
- Cannot remove a view in postges. (Neither DBI nor RPostgres has a method for dropping a view. It must be crafted by the user.)
- Additional options, which must be called as named ones, include
	- `temoprary = FALSE`: a logical specifying whether to consider only temprary tables. **NOT available with postgres**.
	- `fail_if_missing = TRUE`: a logical specifying whether to raise an error for the call to remove an absent table.

```{r remove-example1, collapse=TRUE}
if (con %>% dbExistsTable("mtcars") %>% isFALSE()) {
    con %>% dbWriteTable("mtcars", mtcars)
}

con %>% dbListTables()
con %>% dbRemoveTable("mtcars") # remove here
con %>% dbListTables() # removed table no longer exists
```

In Postgres, `temporary = TRUE` causes an error regardless of the existanece of the table.

```{r remove-temoprary-fails, error=TRUE, collapse=TRUE}
if (con %>% dbExistsTable("mtcars") %>% isFALSE()) {
    con %>% dbWriteTable("mtcars", mtcars)
}
con %>% dbRemoveTable("mtcars", temporary = TRUE) # error
```

In this case, the asked removal will not be executed.

```{r remove-temoprary-fails2, collapse=TRUE}
con %>% dbListTables() # the table is still alive
```

The following examples illustate the behaviour of the option `fail_if_missing =`.

```{r remove-absent-table1, collapse=TRUE}
# fail_if_missing = FALSE makes an otherwise invalid call valid
con %>% dbRemoveTable("absent_table", fail_if_missing = FALSE)
```
```{r remove-absent-table2, error=TRUE, collapse=TRUE}
# fail_if_missing = TRUE, the default, raises an error for deleting an absent table
con %>% dbRemoveTable("absent_table")
```

### `dbReadTable()`

- Copy data frames from database tables.
- Returns a data frame that contains the complete data from the remote table.

```{r }
if (con %>% dbExistsTable("mtcars") %>% isFALSE()) {
    con %>% dbWriteTable("mtcars", mtcars)
}
con %>%
    dbReadTable("mtcars") %>%
    head()
```

### `dbAppendTable()`

- Inserts rows into a table, assuming that the table has been created beforehand.
- Returns a scalar numeric equal to the number of rows inserted.
- The `row.names` argument must be `NULL`, the default value. The row names of input data are ignored.
- In Postgres, `DBI::dbAppendTable()` is overrideen by `RPostgres::dbAppendTable()` because RPostgres uses placeholders of the form `$1`, `$2` etc. instead of `?`.

```{r append-table-mtcars2, collapse=TRUE}
# create demo table
con %>% dbWriteTable("mtcars2", mtcars %>% head(2), overwrite = TRUE)
# insert 10 rows from tail
con %>% DBI::dbAppendTable(name = "mtcars2", value = mtcars %>% tail(10))
# show resulting table
con %>% dbReadTable("mtcars2")
```

Appending rows into an absent table raises an error.

```{r append-absent-table, error=TRUE}
con %>% dbAppendTable("iris", iris %>% head())
```

```{r, include=FALSE}
remove_all_table(con)
con %>% DBI::dbDisconnect()
```