# Views

```{r knit-setting-for-views, eval = TRUE, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = TRUE, warning = FALSE, collapse = TRUE)
```

```{r function-drop-all-views, include=FALSE}
drop_all_views <- function(con) {
    con %>%
        dbListViews() %>%
        purrr::map_int(~ dbDropView(con, .x, cascade = TRUE, if_exists = TRUE)) %>%
        sum()
}
```

```{r function-drop-all-tables, include=FALSE}
drop_all_tables <- function(conn, names.except = NULL, if_exists = TRUE, cascade = TRUE) {
    if (names.except %>% inherits(c("character", "NULL")) %>% isFALSE()) {
        stop("proviede names.except must be of character-like class.")
    }
    con %>%
        DBI::dbListTables() %>%
        base::setdiff(con %>% dbListViews()) %>%
        base::setdiff(names.except) %>%
        purrr::map_int(~ dbDropTables(con, .x, if_exists, cascade)) %>%
        invisible()
}
```

```{r execute-drop-all-views, eval=FALSE, include=FALSE}
con %>% drop_all_views()
```

```{r execute-clear-all-tables, include=FALSE}
# clear views and table, if any
con %>% drop_all_tables()
```

This section asssumes that `mtcars` table has been created in a remote table.

```{r prepare-mtcars-table}
if (con %>% dbExistsTable("mtcars") %>% isFALSE()) {
    con %>% dbWriteTable("mtcars", mtcars)
}

con %>%
    dbReadTable("mtcars") %>%
    head()
```

## Create a View

- Creates a view from the passed SQL query.
- Returns `TRUE` invisibly.
- `temporary = FALSE` and `or_replace = FALSE` arguments specify whether to enable `TEMPORARY` and `OR REPLACE` options, respectively. 

### Definition

```{r def-create-view-dbi-show, ref.label='def-create-view-dbi', eval=FALSE}
```

Since `DBI::dbExecute` does not accept multiple statements, `dbCreateView` is not subject to SQL injection of that form.

### Example

```{r create-view-actually-creates-view-example}
if (con %>% dbExistsTable("cars") %>% isFALSE()) con %>% dbWriteTable("cars", cars)

con %>% dbCreateView(
    name = "view_cars_20",
    sql = "SELECT * FROM cars WHERE dist < 20"
)
con %>% dbReadTable("view_cars_20")

# see the view properly follow the change in the referred table
con %>% dbAppendTable("cars", data.frame(speed = 7, dist = 19))
con %>% dbReadTable("view_cars_20")
```

```{r, include=FALSE}
con %>% drop_all_tables(names.except = "mtcars")
```

temporary views vanish when temporary resources are freed, which happenes, for instance, when `dbDisconnect()` is executed. 

```{r create-temp-view-example}
sql <- "SELECT * FROM mtcars"
con %>% dbCreateView("view_permanent", sql)
con %>% dbCreateView("view_temporary", sql, temporary = TRUE)
con %>% dbListTables()
# reconnect
con %>% dbDisconnect()
con <- issue_con("demo")
# temporary table has vanished
con %>% dbListTables()
```

```{r ref.label='execute-drop-all-views', include=FALSE}
```

Given a set of `SELECT` queries and names that should be complied as views, you can create those views at once.

```{r create-many-views-at-once}
cyls <- c(4, 6, 8)
names <- purrr::map(cyls, ~ paste("view_cyl=", .x, sep = ""))
sqls <- purrr::map(cyls, ~ sqlInterpolate(
    con,
    "SELECT * FROM mtcars WHERE cyl = ?cyl",
    cyl = .x
))
# vectorised view creation
purrr::map2_lgl(names, sqls, ~ dbCreateView(con, .x, .y))
con %>% dbListTables()
```

```{r ref.label='execute-drop-all-views', include=FALSE}
```

The combination of `dbCreateView()` and the dbplyr package allows you to create views without writing any SQL statements. This is often much quicker than working with SQL.

```{r create-views-lazy}
if (con %>% dbExistsTable("flights") %>% isFALSE()) con %>% dbWriteTable("flights", nycflights13::flights)

# work on lazy table
# get all rows with no missing values
flights_no_NA <- tbl(con, "flights") %>%
    filter(if_all(everything(), ~ !is.na(.)))
flights_no_NA %>% class()
flights_no_NA %>% summarise(n())

# number of NA on each columns
flights_no_NA %>% summarise(
    across(
        .cols = everything(),
        .fns = ~ if_else(is.na(.x), 1, 0) %>% sum(),
        .names = "{.col}_NA"
    )
)

# transform it into a view
con %>% dbCreateView(
    name = "flights_no_NA",
    sql = flights_no_NA %>% dbplyr::sql_render()
)

con %>% dbListTables()
flights_view <- con %>% tbl("flights_no_NA")
flights_no_NA %>% summarise(n())
flights_view %>% head()
```

```{r ref.label='execute-drop-all-views', include=FALSE}
```

```{r remove-flights, include=FALSE}
con %>% drop_all_tables(names.except = "mtcars")
```

## List views

- Lists all view names, including temporary ones.
- Returns a vector of strings representing view names.
- The arguments `permanent = TRUE` and `temporary = TRUE` specify whether to include non-temporal and temporary views, respectively. An empty character will be returned if both are set to `FALSE`.

### Definition
```{r def-list-views-show, ref.label='def-list-views'}
```

### Example

```{r list-view-example}
N <- 3
sql <- "SELECT * FROM mtcars"
names <- purrr::map(1:N, ~ paste("view_", .x, sep = ""))
names_tmp <- purrr::map(1:N, ~ paste("view_temp_", .x, sep = ""))
# non-temporary views
purrr::map_lgl(names, ~ dbCreateView(con, .x, sql))
# temporary views
purrr::map_lgl(names_tmp, ~ dbCreateView(con, .x, sql, temporary = TRUE))

# show temoporary views
con %>% dbListViews(permanent = FALSE)
# show non-temoporary views
con %>% dbListViews(temporary = FALSE)
```

```{r ref.label='execute-drop-all-views', include=FALSE}
```

## Test existence of a view

- Test the existence of views.
- Returns a logical vector of the same length as the input.
- Accpepts the same arguments as `dbCreateView`.

### Definition

```{r def-exists-view-show, ref.label='def-exists-view'}
```

### Example

```{r test-views-example}
sql <- "SELECT * FROM mtcars"
purrr::map2_lgl(
    .x = c("view", "view_temp"),
    .y = c(FALSE, TRUE),
    .f = ~ dbCreateView(con, .x, sql, temporary = .y)
)

viewnames <- c("view", "view_temp", "view_absent")
con %>% dbExistsViews(viewnames)
con %>% dbExistsViews(viewnames, permanent = FALSE)
con %>% dbExistsViews(viewnames, temporary = FALSE)
```

```{r ref.label='execute-drop-all-views', include=FALSE}
```

## Drop a view {#dropview}

- Drops a view.
- Returns the total number of views dropped, invisibly.
- The arguments `if_exists = FALSE` and `cascade = FALSE` specify whether to enable the corresponding SQL options. If `cascade = FALSE`, the default, then `RESTRICT` option is set.

### Definition

```{r def-drop-view-show, ref.label='def-drop-view'}
```

### Example

View can be dropped in a vectorised way. If you are dropping temporary views only, then you can also use the `DISCARD` statement \@ref(discard).

```{r drop-all-temporary-views-example}
# create views to demonstrate a vectorised removal
sql <- "SELECT * FROM mtcars LIMIT 10"
names <- purrr::map(1:4, ~ paste("view", .x, sep = ""))
tmps <- purrr::map(1:4, ~ .x <= 2)
purrr::map2_lgl(names, tmps, ~ dbCreateView(con, .x, sql, temporary = .y)) %>% invisible() # discard output

con %>% dbListViews()

# drop only temporary views
con %>%
    dbListViews(temporary = TRUE, permanent = FALSE) %>%
    purrr::map_int(~ dbDropView(con, .x)) %>%
    sum() # total number of tables dropped

con %>% dbListViews() # non-temporary views are still alive

# drop them
con %>%
    dbListViews(temporary = FALSE, permanent = TRUE) %>%
    purrr::map_int(~ dbDropView(con, .x)) %>%
    sum()

con %>% dbListViews()
```

Dependent views could be a source of an error during this vectorised removal. Set `if_exists = TRUE` whenever `cascade = TRUE` to avoid this problem.

```{r drop-view-example-cascading, error=TRUE}
con %>% dbCreateView(
    name = "view_super",
    sql = "SELECT * FROM mtcars LIMIT 10"
)
con %>% dbCreateView(
    name = "view_sub",
    sql = "SELECT * FROM view_super LIMIT 10"
)

# set `if_exists = TRUE` to skip an cascade-dropped view
c("view_super", "view_sub") %>%
    purrr::map(~ dbDropView(con, .x, cascade = TRUE))
```

