```{r knit-setting-for-misc, eval = TRUE, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = TRUE, warning = FALSE, collapse = TRUE)
```

# Misc

## Free Resources {#discard}

- Executes a `DISCARD` SQL statement.
- Returns `TRUE` invisibly even if it does nothing.
- The arguments `all = FALSE`, `plans = FALSE` and `temporary = TRUE` specify which of the corresponding SQL statement option should be applied. If multiple options are selected, only the most dominant one will be executed. If none is selected, nothing will be done.

### Definition

```{r def-discard-function-show, ref.label='def-discard-function'}
```

### Example

`DISCARD TEMPORARY` frees all tempoary resources at once, including tempoary views, instead of listing and dropping them, as we did in \@ref(dropview).

```{r discard-temprary-view}
# create temporary and non-temporary views
sql <- "SELECT * FROM mtcars LIMIT 10"
purrr::map_lgl(
    .x = 1:4,
    ~ dbCreateView(
        con,
        name = paste("view", .x, sep = ""),
        sql = sql,
        temporary = isTRUE(.x <= 2)
    )
)

con %>% dbListViews()
con %>% dbDiscard(temporary = TRUE) # delete all temporary objects
con %>% dbListViews() # non-temporary views are still alive
```

```{r ref.label='execute-drop-all-views', include=FALSE}
```

## Drop a Table {#droptable}

### Definition

- Drop a table, with `IF EXISTS` and `CASCADE` option.
- Returns the total number of views dropped, invisibly.
- See \@ref(dropview) for the usage of optional arguments.

```{r ref.label='def-drop-table', eval=FALSE}
```

### Example

```{r, drop-tables-example, error=TRUE}
sql <- "SELECT * FROM mtcars"
purrr::map_lgl(
    .x = 1:3,
    .f = ~ dbCreateView(con, name = paste0("view", .x), sql)
)
con %>% dbListTables()

# specifying a view causes an error
con %>% dbDropTables("view2", if_exists = TRUE)
con %>% dbListTables()

# specifying a referred table deletes all dependent views
con %>% dbDropTables("mtcars", cascade = TRUE)
con %>% dbListTables()
```

```{r ref.label='execute-clear-all-tables', include=FALSE}
```
```{r exit-disconnect,include=FALSE}
con %>% dbDisconnect()
```