---
title: "Code Examples for DBI and RPostgres"
author: "Shena"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
github-repo: 'Shena4746/datascience-100knocks-preprocess-R'
---

```{r setup, eval = TRUE, include=FALSE}
rm(list = ls())
gc(reset = TRUE)
gc(reset = TRUE)
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, warning = FALSE)
```

# About DBI and RPostgres

```{r libraries, message=FALSE, warning=FALSE}
library("RPostgres")
library("DBI")
library("dbplyr")
library("dplyr")
```

This document is based on the following conditions.
```{r versions-of-libraries, collapse=TRUE}
R.version.string
packageVersion("RPostgres")
packageVersion("DBI")
packageVersion("dbplyr")
packageVersion("dplyr")
packageVersion("purrr")
```

## Overview

## List of Functions

### Functions in DBI 

```{r DBI-functions, collapse=TRUE}
# functions in dbi package
dbi <- ls("package:DBI")
dbi %>% length() # how many ?
dbi %>% print()
```

### Functions in RPostgres

```{r RPostgres-functions, collapse=TRUE}
# functions in RPostgres package
rpos <- ls("package:RPostgres")
rpos %>% length() # how many ?
rpos %>% print()
```

### Functions with duplicated names

You might have noticed that there are quite a few functions with the same name in these packages. They provides the same functionlaity, possibly with the different implementation and efficiecy.

```{r DBI-RPostgres-common-functions, collapse=TRUE}
# common functions in DBI and RPostgres
common <- base::intersect(dbi, rpos)
common %>% length() # how many ?
common %>% print()
```

We don't go into the (potentially deep) difference of these twin functions. This document prefers to using `DBI::` methods as introduced in [GitHub repository of RPostgres package](https://github.com/r-dbi/RPostgres), unless there is a significant difference between them. 

### Funcions only in DBI

```{r DBI-unique-functions, collapse=TRUE}
# functions only in DBI
only_in_dbi <- base::setdiff(dbi, rpos)
only_in_dbi %>% length() # how many ?
only_in_dbi %>% print()
```

### Funcions only in RPostgres

```{r RPostgres-unique-functions, collapse=TRUE}
# functions only in RPostgres
only_in_rpos <- base::setdiff(rpos, dbi)
only_in_rpos %>% length() # how many ?
only_in_rpos %>% print()
```

## Basic Usage

Before we look at indivisual functions, let us describe a typical work flow with these packages.
You may sefaly skip this section as it is independent of what follows.

This section assumes that you have access to a postgres database. If necessary, run the following command to create a new database named 'demo'.

```{bash start-postgres, eval=FALSE}
sudo service postgresql start
```

```{bash, eval=FALSE}
createdb -p 5432 --owner=your-owner-name --host=localhost --username=your-username demo
```

The following code defines the `con` variable which keeps the establised connection to your database. 

```{r example-connect, eval = FALSE}
con <- DBI::dbConnect(
    drv = RPostgres::Postgres(),
    host = "localhost",
    port = 5432,
    dbname = "your-database-name",
    user = "your-user",
    password = "your-password"
)
```

You should not write this way outside this demo. See \@ref(dbConnect.)

```{r example-issue-con, include=FALSE}
dw <- config::get("demo")
con <- DBI::dbConnect(
    drv = RPostgres::Postgres(),
    host = dw$host,
    port = dw$port,
    dbname = dw$dbname,
    user = dw$user,
    password = dw$password
)
```

Let us observe the list of tables in the connected database.

```{r empty-database, collapse=TRUE}
# demo database is empty
con %>% DBI::dbListTables()
```

`dbWriteTable()` Creates a remote table from local data.
The last two arguments are optional.

```{r our-first-table, collapse=TRUE}
# create table by supplying local dataframe
con %>% dbWriteTable("mtcars", mtcars, overwrite = TRUE, temoprary = TRUE)
con %>% dbWriteTable("iris", iris, overwrite = TRUE, temoprary = TRUE)
# check that tables are properly created
DBI::dbListTables(con)
```

You can also ask the existance of the specific table in the database associated with `con` variable.

```{r , collapse=TRUE}
# boolean question about the existance of indivisual table
con %>% DBI::dbExistsTable("mtcars")
```

`dbListFields()` gets you the list of fields of an indivisual table.

```{r list-fields-mtcars, collapse=TRUE}
# now we have
con %>% DBI::dbListFields("mtcars")
```

`dbReadTable()` allows you to take a quick peep at a table. 

```{r read-table-without-query}
con %>%
    DBI::dbReadTable("mtcars") %>%
    head()
```


You can use `dbSendQuery()` to execute a query written in a string. Note that we don't need a `;` at the end of the query. Since this method does nothing but sending query, you should capture the returned value to perform an appropriate post-processing.

```{r first-send-query}
# send query and get response variable
rs <- con %>% DBI::dbSendQuery("
	SELECT *
	FROM mtcars
	WHERE hp <= 100
")
```

`dbFetch()` shows you the result of the query.

```{r first-fetch}
rs %>% DBI::dbFetch()
```

Don't forget to do `dbClearResult()`. 
DBI package keeps many kinds of resources associated with the result of the query unless manually cleared. `dbClearResult()` is responsible for that freeing role.
DBI reference says that 

> This step is mandatory for all objects obtained by calling dbSendQuery() or dbSendStatement().

```{r first-clear-result}
rs %>% dbClearResult()
```

We can write a parameterised query in the following way.
`$1` and `$2` corresponds to the first and second argument in the `list(100, 120)`, respectively. 

```{r first-param-query, collapse=TRUE}
rs <- con %>% dbSendQuery("
	SELECT *
	FROM mtcars
	WHERE hp BETWEEN $1 AND $2
")
rs %>% dbBind(list(100, 120))
rs %>% dbFetch()
rs %>% dbClearResult()
```

You can delete a table by `dbRemoveTable()`. 

```{r remove-tables}
# remove indivisual table by supplying name string
con %>% DBI::dbRemoveTable("mtcars")
```

It is often useful to have a function that deletes multiple tables at once based on a pre-defined rule.
The following code is one of the most simplest, deleting all tables.

```{r, collapse=TRUE}
# create mtcars table again to demonstate sumultaneous-removal
con %>% DBI::dbWriteTable("mtcars", mtcars, overwrite = TRUE, temoprary = TRUE)
DBI::dbListTables(con)
# remove multiple tables at once
con %>%
    DBI::dbListTables() %>%
    purrr::map_lgl(~ DBI::dbRemoveTable(con, .x)) %>%
    all()
DBI::dbListTables(con)
```

`dbDisconnect()` terminates the supplied `con` variable, and closes the connection.

```{r first-dbDisconnect}
con %>% DBI::dbDisconnect()
```